### 2018-05-02 papers/synth/prelim.tex
as

\begin{df}
  A VTS $G$ is a tuple $(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$, where
  \begin{itemize}
  \item $\nodes$ is a set of nodes representing compartments in the VTS,
  \item $\mols$ is the set of molecules flowing in the system, 
  \item $\edges \subseteq \nodes \times (\powerset{\mols}-\emptyset) \times \nodes$ is the
    set of edges with molecule sets as labels,
  \item $\nlabel : \nodes \maps \powerset{\mols}$ defines the molecules present in the nodes,
  \item $\pairs \subseteq \mols \times \mols$ is pairing relation,
  \item $\nodef : \mols \maps \powerset{\mols} \maps \booleans$ is activity maps for nodes, and
  \item $\edgef : \mols \maps \powerset{\mols} \maps \booleans $ is activity maps for edges.
  \end{itemize}
\end{df}
$\nodes$, $\mols$, $\edges$, and $\nlabel$ define a labelled graph.
%
Additionally, $\pairs$ defines which molecules can fuse with which molecules,
and
$\nodef$ and $\edgef$ are the activity functions for molecules on
nodes and edges respectively.
%
The model captures the study state of a VTS.
%
The analysis of the model will inform us about the network/graph
properties of VTSs.

Let $\pairs(M')$ denote $\{m|(m,m') \in P \text{ and } m' \in M'\}$.
%
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and
$\nodef(k,\nlabel(n))$ is true.
%
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and
$\edgef(k,M')$ is true.
%
We call $G$ {\em well-structured} if molecules $M$ is divided into
two equal-sized partitions $M_1$ and $M_2$ such that
$P \subseteq M_1 \times M_2$, and
for each $(n,M',n') \in \edges$, $n \neq n'$ and
$M' \subseteq \nlabel(n) \intersection \nlabel(n')$.
%
In other words,
no molecule can participate in fusion from both channels and
compartments,
%
there are no self loops, and 
each edge carry only those molecules that are present in its source
and destination nodes. 
%
An edge $(n,M',n') \in \edges$ {\em fuses} with its destination node $n'$
if there are molecules $m,m' \in \mols$ such that $m$ is active in
$(n,M',n')$, $m'$ is active in $n'$, and $(m,m') \in \pairs$.
%
We call $G$ {\em well-fused} if each edge $(n,M',n') \in \edges$ fuses
with non-empty fusing molecules $M'' \subseteq M'$
and $\pairs(M'')$ are not active in any other node.

@@

\item Boolean variable $n_{i,m}$ indicates if $m \in \nlabel(i)$
\item Boolean variable $e_{i,j,q}$ indicates if $q$th edge exists between $i$ and $j$.
\item Boolean variable $e_{i,j,q,m}$ indicates if $q$th edge between $i$ and $j$ contains $m$.
\item Boolean variable $p_{m,m'}$ indicates if $(m,m') \in \pairs$
\item uninterpreted Boolean functions $\nodef_m : \booleans^\mu \maps \booleans$
encoding $\nodef(m)$ map
\item uninterpreted Boolean functions $\edgef_m : \booleans^\mu \maps \booleans$
encoding $\edgef(m)$ map
\end{enumerate}
We also have auxiliary Boolean variables that will help us encode the well-fused property. 
\begin{enumerate}
\item $a_{i,m}$ indicates that molecule $m$ is active at node $i$, i.e., $\nodef(m,L(i))$
  holds
\item $b_{i,j,q,m}$ indicates that molecule $m$ is active at $q$th edge $(i,M',j)$ between $i$ and $j$, i.e., $\edgef(m,M')$ holds
\end{enumerate}

@@

The following constraints encode the basic consistency of VTSs.
\begin{align*}
  \texttt{EdgeC} =\;& \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
  \land
  \bigwedge\limits_{i,q} \neg e_{i,i,q}
  \land
  % \tag{\texttt{NodeC}}\label{eq:f0}\\
  \bigwedge\limits_{\mathclap{i,j,q,m}} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
  \\
  \texttt{ActivityC} =\;&
  \bigwedge\limits_{\mathclap{i,j,q,m}} b_{i,j,q,m} \limplies e_{i,j,q,m} \quad\land\quad
  \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
  % \tag{V2}\label{eq:f1}\\
  \\
  \texttt{PairingC} =\;&
  \bigwedge\limits_{m} 
  (\Lor\limits_{m'} p_{mm'} \limplies \Land_{m'} \lnot p_{m'm} ) \land
  (\Lor\limits_{m'} p_{m'm} \limplies \Land_{m'} \lnot p_{mm'} )
  \\
  \texttt{Fusion1} =\;&
  \bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
  \bigvee_{\mathclap{m,m^{\prime}}} (b_{i,j,q,m} \land a_{j,m^{\prime}} \land p_{m,m^{\prime}})
  \\
  \texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m}} b_{i,j,q,m} \limplies \neg 
  \bigvee_{\mathclap{j \neq j^{\prime}, m^{\prime\prime}}} ( a_{j^{\prime},m^{\prime\prime}} \land p_{m,m^{\prime\prime}})
  \\
  \texttt{Consistancy} =\;& \texttt{EdgeC} \land
  \texttt{ActivityC} \land \texttt{PairingC} \land
  \texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
\texttt{EdgeC} states that each edge has at least one molecule,
there are no self loops, and edge labels are consistent with node labels.   
\texttt{ActivityC} states that active molecule are present.
\texttt{PairingC} states that a molecule can either participate in fusion
via nodes or via edges, but not both.
\texttt{Fusion1}, and \texttt{Fusion2} states the well-fused condition.
\texttt{Consistancy} is the conjunction of all of the above.

>>>>

\begin{df}
  A VTS $G$ is a tuple $(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$, where
  \begin{itemize}
  \item $\nodes$ is a set of nodes representing compartments in the VTS,
  \item $\mols$ is the set of molecules flowing in the system, 
  \item $\edges \subseteq \nodes \times (\powerset{\mols}-\emptyset) \times \nodes$ is the
    set of edges with molecule sets as labels,
  \item $\nlabel : \nodes \maps \powerset{\mols}$ defines the molecules present in the nodes,
  \item $\pairs \subseteq 2^{\mols}$ is pairing relation,
  \item $\nodef : \mols \maps \powerset{\mols} \maps \booleans$ is activity maps for nodes, and
  \item $\edgef : \mols \maps \powerset{\mols} \maps \booleans $ is activity maps for edges.
  \end{itemize}
\end{df}
$\nodes$, $\mols$, $\edges$, and $\nlabel$ define a labelled graph.
%
Additionally, $\pairs$ defines which molecules can fuse with which molecules,
and
$\nodef$ and $\edgef$ are the activity functions for molecules on
nodes and edges respectively.
%
The model captures the study state of a VTS.
%
The analysis of the model will inform us about the network/graph
properties of VTSs.

% Let $\pairs(M')$ denote $\{P\setminus M'|  P \in \pairs \text{ and } P \intersect M' \neq \emptyset \}$.
%
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and
$\nodef(k,\nlabel(n))$ is true.
%
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and
$\edgef(k,M')$ is true.
%
We call $G$ {\em well-structured} if molecules $M$ is divided into
two partitions $Q$ and $R$ such that
for each $P \in \pairs, |P \intersect Q| \leq 3 \land |P \intersect R| = 1 $, and
for each $(n,M',n') \in \edges$, $n \neq n'$ and
$M' \subseteq \nlabel(n) \intersection \nlabel(n')$.
%
In other words,
no molecule can participate in fusion from both channels and
compartments,
%
there are no self loops, and 
each edge carry only those molecules that are present in its source
and destination nodes. 
%
An edge $(n,M',\_) \in \edges$ {\em fuses} with a node $n'$
if there are non-empty set of molecules $M'' \subseteq M'$ and $M''' \subseteq \nlabel(n')$
such that $M''$ are active in the edge, $M'''$ are active in $n'$, and $M'' \union M''' \in \pairs$.
%
We call $G$ {\em well-fused} if each edge $(n,M',n') \in \edges$ fuses
with its destination node $n'$
and can not fuse with any other node.

% $\pairs(M'')$ are not active in any other node.

@@

\item Boolean variable $n_{i,m}$ indicates if $m \in \nlabel(i)$
\item Boolean variable $e_{i,j,q}$ indicates if $q$th edge exists between $i$ and $j$.
\item Boolean variable $e_{i,j,q,m}$ indicates if $q$th edge between $i$ and $j$ contains $m$.
\item Boolean variable $p_{m_1,m_2,m_3,m'}$ indicates if $\{m_1,m_2,m_3,m'\} \in \pairs$
\item uninterpreted Boolean functions $\nodef_m : \booleans^\mu \maps \booleans$
encoding $\nodef(m)$ map
\item uninterpreted Boolean functions $\edgef_m : \booleans^\mu \maps \booleans$
encoding $\edgef(m)$ map
\end{enumerate}
We also have auxiliary Boolean variables that will help us encode the well-fused property. 
\begin{enumerate}
\item $a_{i,m}$ indicates that molecule $m$ is active at node $i$, i.e., $\nodef(m,L(i))$
  holds
\item $b_{i,j,q,m}$ indicates that molecule $m$ is active at $q$th edge $(i,M',j)$ between $i$ and $j$, i.e., $\edgef(m,M')$ holds
\end{enumerate}

@@

The following constraints encode the basic consistency of VTSs.
\begin{align*}
  \texttt{EdgeC} =\;& \bigwedge\limits_{i,j,q} (\bigvee_m e_{i,j,q,m} )\limplies e_{i,j,q}
  \land
  \bigwedge\limits_{i,q} \neg e_{i,i,q}
  \land
  % \tag{\texttt{NodeC}}\label{eq:f0}\\
  \bigwedge\limits_{\mathclap{i,j,q,m}} e_{i,j,q,m} \limplies (n_{i,m} \land n_{j,m} )
  \\
  \texttt{ActivityC} =\;&
  \bigwedge\limits_{\mathclap{i,j,q,m}} b_{i,j,q,m} \limplies e_{i,j,q,m} \quad\land\quad
  \bigwedge\limits_{i,m} a_{i,m} \limplies n_{i,m}
  % \tag{V2}\label{eq:f1}\\
  \\
  \texttt{PairingC} =\;&
  \exists qr. \Land\limits_{\mathrlap{m_1,m_2,m_3,m_4}\;\;}
                         (p_{m_1,m_2,m_3,m_4} \limplies qr_{m_1} + qr_{m_2} + qr_{m_3} + qr_{m_4} = 3) \land\\ 
  & \quad\quad\quad ( m_1 = m_2 = m_3 = m_4  \limplies \lnot p_{m_1,m_2,m_3,m_4} )
  % \bigwedge\limits_{m} 
  % (\Lor\limits_{m'} p_{mm'} \limplies \Land_{m'} \lnot p_{m'm} ) \land
  % (\Lor\limits_{m'} p_{m'm} \limplies \Land_{m'} \lnot p_{mm'} )
  \\
  \texttt{Fusion1} =\;&
  \bigwedge\limits_{i,j,q} e_{i,j,q} \limplies
  \bigvee_{{m_1,m_2,m_3,m_4}} (\Land_{l=1}^4 ( b_{i,j,q,m_l} \lor a_{j,m_l} ) \land 
                        \sum_{l=1}^4 b_{i,j,q,m_l} > 0 \land\\
  & \qquad \qquad \qquad \qquad \sum_{l=1}^4 a_{j,m_l} > 0 \land p_{m_1,m_2,m_3,m_4})
  \\
  \texttt{Fusion2} =\;&
\bigwedge\limits_{\mathclap{i,j,q,m_1,m_2,m_3, l \in \{1,..,3\}}} b_{i,j,q,m_1} \land ..\land b_{i,j,q,m_l} \limplies \neg 
  \bigvee_{\mathclap{j \neq j^{\prime}, m_{l+1}^{\prime},..,m_{4}^{\prime}}} ( a_{j^{\prime},m_{l+1}^{\prime}} \land .. \land a_{j^{\prime},m_4^{\prime}} \land p_{m_1,..,m_l,m^{\prime}_{l+1},..,m^{\prime}_4})
  \\
  \texttt{Consistancy} =\;& \texttt{EdgeC} \land
  \texttt{ActivityC} \land \texttt{PairingC} \land
  \texttt{Fusion1} \land \texttt{Fusion2} 
\end{align*}
\texttt{EdgeC} states that each edge has at least one molecule,
there are no self loops, and edge labels are consistent with node labels.   
\texttt{ActivityC} states that active molecule are present.
\texttt{PairingC} states that a molecule can either participate in fusion
via nodes or via edges, but not both.
\texttt{Fusion1}, and \texttt{Fusion2} states the well-fused condition.
\texttt{Consistancy} is the conjunction of all of the above.
